---
title: "Reproducible Data Analysis: Airports, Monte Carlo Integration, and GenAI Prompting"
author: "Akshith Kachakayala"
date: today
date-format: "MMMM D, YYYY"
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    colorlinks: true
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
    fontsize: 11pt
    include-in-header:
      text: |
        \usepackage{float}
        \floatplacement{figure}{H}
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# ── Package loading ──────────────────────────────────────────────────────────
library(rvest)       # web scraping
library(tidyverse)   # data wrangling + ggplot2
library(knitr)       # kable()
library(kableExtra)  # table formatting
```

# Busiest Airports Analysis

## Overview

This section analyzes passenger traffic at six of the world's busiest airports
from 2020 through 2025. Data are scraped directly from Wikipedia's
*List of busiest airports by passenger traffic*, ensuring the analysis is
reproducible and current. The six airports examined are Atlanta Hartsfield–Jackson
(ATL), Frankfurt (FRA), Beijing Daxing (PKX), Phoenix Sky Harbor (PHX),
Chicago O'Hare (ORD), and London Heathrow (LHR).

## Data Collection and Preparation

```{r}
#| label: airports-scrape

## Step 2a: Scrape the data; 2020-2025
airportWikiData <- read_html(
  x = "https://en.wikipedia.org/wiki/List_of_busiest_airports_by_passenger_traffic"
) |>
  html_elements(css = "table") |>
  html_table() |>
  lapply(function(tbl) {
    names(tbl) <- make.names(names(tbl), unique = TRUE)
    tbl
  })

## Step 2b: Extract data and add year columns
airports_2025 <- airportWikiData[[2]] |> mutate(year = 2025, .before = 1)
airports_2024 <- airportWikiData[[3]] |> mutate(year = 2024, .before = 1)
airports_2023 <- airportWikiData[[4]] |> mutate(year = 2023, .before = 1)
airports_2022 <- airportWikiData[[5]] |> mutate(year = 2022, .before = 1)
airports_2021 <- airportWikiData[[6]] |> mutate(year = 2021, .before = 1)
airports_2020 <- airportWikiData[[7]] |> mutate(year = 2020, .before = 1)
```

```{r}
#| label: airports-wrangle

## Step 2c: Combine data frames & filter to target airports
## After setNames + make.names, special-character column names are sanitized:
##   "%change"         -> "X.change"
##   "Code(IATA/ICAO)" -> "Code.IATA.ICAO."
##   "Totalpassengers" -> unchanged

targetAirports <- c("ATL", "FRA", "PKX", "PHX", "ORD", "LHR")

airportData6 <- bind_rows(
  airports_2020,
  airports_2021,
  airports_2022,
  airports_2023,
  airports_2024,
  airports_2025
) |>
  rename_with(~ "passengers",
              .cols = matches("otalpassenger", ignore.case = TRUE)) |>
  rename_with(~ "percentChange",
              .cols = matches("^X\\.change$|^X\\.Change$|\\.change$",
                              ignore.case = TRUE)) |>
  rename_with(~ "codeRaw",
              .cols = matches("IATA", ignore.case = TRUE)) |>
  separate_wider_delim(
    cols     = codeRaw,
    delim    = "/",
    names    = c("IATA", "ICAO"),
    too_few  = "align_start",
    too_many = "merge"
  ) |>
  filter(IATA %in% targetAirports) |>
  mutate(
    across(
      .cols = c(passengers, percentChange),
      .fns  = ~parse_number(.x)
    )
  )

## Clean up excess data tables
remove(
  list = c("airports_2020", "airports_2021", "airports_2022",
           "airports_2023", "airports_2024", "airports_2025",
           "airportWikiData")
)
```

Two data visualizations of a table and a plot were created that compare six of the busiest passenger airports in the world over a six year time period (2020 to 2025). The six airports are Hartsfield-Jackson Atlanta International Airport (ATL), Frankfurt Airport (FRA), Beijing Daxing International Airport (PKX), Dubai International Airport (DXB), London Heathrow Airport (LHR), and Tokyo Haneda Airport (HND)



## Summary Table

The plot displays a line graph with year on the horizontal axis running from 2020 to 2025 and total passengers in millions on the vertical axis starting at zero. All the airports saw a drop in passengers due to the COVID 19 pandemic in the years 2020 and 2021. Atlanta and Dubai had the highest passenger totals in the later years, while Beijing Daxing had a slower rebound after the pandemic.

```{r}
#| label: tbl-airports
#| tbl-cap: "Total annual passengers (2020–2025) for six of the world's busiest airports."

# Pivot to wide format: one row per airport, one column per year
airportData6_wide <- airportData6 |>
  dplyr::select(year, Airport, IATA, passengers) |>
  pivot_wider(
    names_from  = year,
    values_from = passengers
  )

# Render polished table
options(knitr.kable.NA = "")
airportData6_wide |>
  # Clean any Wikipedia citation brackets from airport names (e.g. "Atlanta[3]")
  mutate(Airport = str_remove(Airport, "\\[.*?\\]")) |>
  rename(`Airport Name` = Airport) |>
  kable(
    align       = "llccccccc",
    format.args = list(big.mark = ",")
  ) |>
  kable_classic() |>
  row_spec(
    row        = c(1, 5, 6),
    background = "#FFFFC5"
  ) |>
  footnote(
    general = "Airports absent from a year's top-50 ranking have missing passenger data. Highlighted rows are the three required airports.",
    footnote_as_chunk = TRUE
  )
```

## Visualization

Looking at both the table and the line plot together, it is clear that all six airports took a huge hit in passenger numbers in 2020 because of the COVID-19 pandemic, but they did not all recover at the same speed. Phoenix and Atlanta bounced back the fastest, which makes sense because those airports handle a lot of domestic flights within the United States, and Americans started traveling again pretty quickly once restrictions lifted. Frankfurt and Tokyo Haneda recovered more slowly because they rely more heavily on international travelers, who took longer to return. China kept its borders closed much longer than the other countries so predictably Beijing has one of the slowest progressions in passenger numbers. By 2024 and 2025 most of the airports had climbed back close to or above their pre-pandemic passenger levels, but depending on the political context and locality, some countries were slower.

```{r}
#| label: fig-airports-line
#| fig-cap: "Annual passenger traffic (in millions) at six major airports, 2020–2025. Lines connect annual observations; point markers show exact values."
#| fig-height: 4.5
#| fig-width: 7.5

# PSU brand color palette
psuPalette <- c("#1E407C", "#BC204B", "#3EA39E", "#E98300",
                "#999999", "#AC8DCE", "#F2665E", "#99CC00")

airportData6 |>
  mutate(passengers_mil = passengers / 1e6) |>
  ggplot(
    mapping = aes(
      x        = year,
      y        = passengers_mil,
      group    = IATA,
      color    = IATA,
      linetype = IATA
    )
  ) +
  geom_point(size = 2) +
  geom_path(linewidth = 0.8) +
  theme_bw() +
  labs(
    x        = "Year",
    y        = "Total Passengers (millions)",
    title    = "Total Annual Passengers 2020–2025 for Six Busy Airports",
    color    = "Airport",
    linetype = "Airport",
    alt      = "Line graph showing passenger trends at six busy airports from 2020 to 2025."
  ) +
  scale_color_manual(values = psuPalette) +
  scale_x_continuous(breaks = 2020:2025) +
  theme(
    legend.position   = "bottom",
    legend.key.width  = unit(2, "cm"),
    legend.key.height = unit(0.25, "cm")
  )
```

---

# Monte Carlo Numerical Integration

## Background

Monte Carlo numerical integration estimates a definite integral geometrically like randomly scattered points uniformly across a bounding rectangle which checks points to see if it falls below or above a curve to estimate the ratio of the curve's area to the rectangle's known area. 


## Assigned Function

The function assigned for this analysis is the **Beta(2, 2) probability density
function**:

$$
f(x) = \frac{x^{1}(1-x)^{1}}{B(2,\,2)} = 6x(1-x), \quad x \in [0,\,1]
$$

The bounding rectangle uses $x \in [0,1]$ and $y \in [0, 1.5]$, giving a box
area of $1.5$. Because $f(x)$ is a valid probability density, the true integral
equals exactly $1$.

## Simulation and Point Generation

```{r}
#| label: mc-packages

library(patchwork)  # combine ggplot objects into small multiples
```

```{r}
#| label: mc-generate-points

# ── Inline definition of generatePoints() ───────────────────────────────────
# Generates n uniform random (x, y) pairs within the specified bounds.
# Equivalent to the sourced MC_GenPoints.R helper used in class.
generatePoints <- function(xbounds, ybounds, n) {
  tibble(
    x = runif(n, min = xbounds[1], max = xbounds[2]),
    y = runif(n, min = ybounds[1], max = ybounds[2])
  )
}

# PSU brand color palette
psuPalette <- c("#1E407C", "#BC204B", "#3EA39E", "#E98300",
                "#999999", "#AC8DCE", "#F2665E", "#99CC00")

# Generate the master point cloud (n = 10,000)
set.seed(184)
myPoints <- generatePoints(
  xbounds = c(0, 1),
  ybounds = c(0, 1.5),
  n       = 10000
)

# Flag each point as above or on/below the Beta(2,2) curve;
# compute running cumulative proportion and numeric integral estimate
myPoints <- myPoints |>
  mutate(
    flag = case_when(
      y > dbeta(x = x, shape1 = 2, shape2 = 2) ~ "above",
      .default = "on/below"
    ),
    numFlag = case_when(                          # 1 = on/below, 0 = above
      y > dbeta(x = x, shape1 = 2, shape2 = 2) ~ 0,
      .default = 1
    ),
    case   = row_number(),
    prop   = cumsum(numFlag) / case,
    numInt = prop * (1 - 0) * (1.5 - 0)          # prop × x-width × y-height
  )
```

```{r}
#| label: mc-build-plots

# ── Helper: build one panel for a given sample size n ────────────────────────
make_mc_plot <- function(pts, n) {
  pts |>
    slice_head(n = n) |>
    ggplot(mapping = aes(x = x, y = y, shape = flag, color = flag)) +
    stat_function(
      fun   = dbeta,
      args  = list(shape1 = 2, shape2 = 2),
      xlim  = c(0, 1),
      color = "black"
    ) +
    geom_point(size = 0.8, alpha = 0.7) +
    scale_x_continuous(limits = c(0, 1),   expand = expansion()) +
    scale_y_continuous(limits = c(0, 1.5), expand = expansion()) +
    scale_color_manual(values = psuPalette) +
    labs(
      title    = "Monte Carlo Numeric Integration",
      subtitle = paste0("Resolution n = ", format(n, big.mark = ",")),
      caption  = paste(
        "Est. Integral:",
        round(pts$numInt[n], digits = 4)
      ),
      x = "x",
      y = "y"
    ) +
    theme_bw() +
    theme(legend.position = "none")
}

# Build each panel by subsetting the master point cloud
plot10    <- make_mc_plot(myPoints, 10)
plot100   <- make_mc_plot(myPoints, 100)
plot1000  <- make_mc_plot(myPoints, 1000)
plot10000 <- make_mc_plot(myPoints, 10000)
```

## Small-Multiple Visualization

Figure 2 displays the Monte Carlo simulation at four resolutions. Each point is
colored by whether it falls above (blue) or on or below (red) the Beta(2, 2) curve.

```{r}
#| label: fig-mc-smallmult
#| fig-cap: "Monte Carlo numerical integration of the Beta(2, 2) PDF at four sample sizes. Points are colored by position relative to the curve; each panel caption gives the area estimate at that resolution."
#| fig-height: 6.5
#| fig-width: 8.5

plot10 + plot100 + plot1000 + plot10000
```
\newpage

# Code Appendix {.unnumbered}

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
